|

### 2025-09-17 output/Response_to_action_editor_reviewers-2.Rmd

###### Then one could make another version of the left panel with contours for the rho=0.5 population distribution. Comparison of that with the previously-described version would show that the arrows representing discrimination do not have the same numerical meaning with respect to the variance (standard deviation) of the population distribution *in the direction of the arrows*, which is part of the point of all of this.

###### Two graphics with orthogonal axes but different population distributions are the way orthogonal vs. oblique solutions are represented in MIRT analyses. The ms. misses an opportunity to clarify that. Different population distributions on orthogonal axes are the way MIRT folks have thought about factor correlation; factor analysts have thought about 'rotation' and oblique axes (the left panel of Figure 1); it's two sides of the same coin.

###### This graphic is potentially confusing. In addition to the absence of axis labels mentioned above, it has oddly placed gray dotted grid lines. I would think it would have vertical gray dotted grid lines showing placement on the horizontal axis (x), and oblique gray dotted grid lines (orthogonal to y') showing placement on the (oblique) axis y'. With axis values -2 ... 0 ... +2 at the margins of *those* gridlines. Instead, it has the opposite: labeled horizontal grid lines that don't correspond to any axis values, and oblique grid lines labeled 0-1-2-3-4 that similarly don't correspond to any axis, but rather to how far away a point is from y'. Basically, the graphic would be easier to understand with labels on the axes, and two sets of grid lines orthogonal to those that are currently there (omitting those that are currently there).

We understand this graphic may be hard to interpret, but the grid lines are properly placed; the rationale is this:
Each point in the test space is intended to represent a coordinate vector, which is the displacement from the origin given by a linear combination of the basis vectors, with the components of the coordinate vector being the coefficients of those basis vectors (please refer to the expression $`r DISCR_COORDS_ALT_EQ`$ in p. 8 of the previous version of our manuscript, or p. <!-- # TODO: Complete --> in the new submission).
For example, the tip of the vector representing item 1 is at coordinate (+2, +2).
This means the point is displaced twice the length of the first vector basis along its direction (i.e., along the horizontal axis), that is, two units along the horizontal axis, and then another two units parallel to the direction of the oblique axis.
These displacements, parallel to the direction of the axes, is what the grid lines intend to represent.
This representation is somewhat equivalent to a _factor pattern_ (i.e., description in linear combination terms, see @harman_modern_1976, p. 267) representation in the factor model, while the one proposed by the reviewer would be somewhat equivalent to the _factor structure_ (i.e., description in terms of correlations with the latent dimensions) representation. Please see @harman_modern_1976 (p. 265 and p. 267, respectively) for the two representations.
This can be more easily understood by comparing figures 4.2 (p. 61, a representation equivalent to the one we propose) and 12.8 (p. 266) in @harman_modern_1976.

### 2025-09-17 output/Letter_to_editor-2.Rmd
Add Review 2 "letter to editor"

June 2, 2025.

@@

Please find attached the updated version of our manuscript
*"A generalized definition of multidimensional item response theory
parameters"*.
We hope that this new version lives up to the expectations of
the action editor and the reviewers;
we have put as much effort as we have been able to address all the comments and
suggestions, which we believe have helped improve the quality of our paper.
You will find our thorough responses in the document attached to this letter.
However, we acknowledge we may not have resolved satisfactorily all the action
editor's and reviewers' concerns.
Therefore, we would kindly ask you to forward to them our good will
and predisposition to address any other comments or suggestions,
or clarify further on the previous ones.

We thank you in advance for considering our reviewed manuscript.
Please feel free to contact us again at my aforementioned email address
(dvmorillo@psi.uned.es).
We will be happy to address any additional comments or suggestions
the action editor or the reviewers may deem necessary.

>>>>

October XX <!-- # TODO: Complete -->, 2025.

@@

Please find attached our second review of the manuscript
*"A generalized definition of multidimensional item response theory
parameters"*.
we have carefully taken into account Thissen's and your own
review and suggestions, addressing all the comments and suggestions diligently.
Where we have decided not to follow a certain suggestion,
we have striven to explain our reasons in the clearest possible way.
Also, besides addressing the comments,
we have added an additional clarification
about the "covariance matrix whitening",
a terminology previously unknown to us.

Once again, we thank you for considering our reviewed manuscript.
Please feel free to contact us again regarding any further concerns at
[dvmorillo@psi.uned.es](mailto:dvmorillo@psi.uned.es).
